### 2021-04-21 submission/manuscript.Rmd
Add to results section

**Five-day post-infection community data can predict mice that will have prolonged *C. difficile* colonization** After identifying

@@

* Discussion of prolonged persistence. C. difficile sequences detected in tissue samples. Association with mucin-degrading bacteria suggested by recent papers.

>>>>

``` {r ml results}
#AUC results
auc_results <-  read_csv(here("results/performance_results.csv")) %>% 
  mutate(taxa_level = "genus") %>%  
  group_by(method) %>% 
  summarize(median = median(AUC)) %>% 
  arrange(desc(median))
#Highest AUC value
max_auc <- auc_results %>% select(median) %>% max()

```
**Five-day post-infection community data can predict mice that will have prolonged *C. difficile* colonization** After identifying bacteria associated with the 5-day, 1-day and post-CDI 1-day PEG treatments, we decided to examine the taxa that were influencing prolonged *C. difficile* colonization. We trained 3 types of machine learning models (random forest, logistic regression, and support vector machine) with input bacterial community data from 5 days post-infection to predict whether the mice were colonized with *C. difficile* 10 days post-infection. We chose 5 days post-infection because that was the earliest timepoint where we would see a treatment effect in the mice that were given 1-day PEG treatment three days after *C. difficile* challenge and then administered an FMT or PBS gavage. The random forest model had the highest performance (AUROC = `r max_auc`, Data Set S1, Sheet X), so we next performed permutation importance to examine the bacteria that were driving performance. We selected the top 20 bacteria contributing to our models performance and examined their relative abundance at 5 days post-infection, the timepoint used to predict C. difficile colonization status on day 10 (Fig. 6A). Next, we focused on the 5 genera that were > 1 % relative abundance in either the cleared or colonized mice and examined the relative abundance dynamics of these bacteria over time. We found _, _, and _ tended to have a higher relative abundance and __, __ had a lower relative abundance in the mice with prolonged colonization compared to the mice that cleared *C. difficile* (Fig. 6B).
Previous work examining the impact of PEG on the murine microbiota found that PEG treatment resulted in the permanent loss of S24-7, also known as *Muribaculum intestinale* [@Tropini2018]. We decided to check our *Porphyromonadaceae* OTUs because *Muribaculum intestinale* is known to be classified as *Porphyromonadaceae* by the Ribosomal Database Project (RDP) database [@Vornhagen2020], *Porphyromonadaceae* was a top feature in the random forest model predicting prolonged *C. difficile* colonization, and had a high relative abundance in the communities of mice that cleared *C. difficile* within 10 days. We identified 4 OTUs that had at least (insert minimum percent identity) to *Muribaculum intestinale* and examined their abundance in mice that either cleared or were still colonized with *C. difficile* at 10 days post-infection (Fig. S3).

@@

* Discussion of prolonged persistence. *C. difficile* sequences detected in tissue samples. Association with mucin-degrading bacteria suggested by recent papers.

### 2021-04-21 submission/manuscript.tex
Add to results section

\textbf{Five-day post-infection community data can predict mice that
will have prolonged \emph{C. difficile} colonization} After identifying

@@

\begin{itemize}
\item
  Summary of major findings (Fig. 7A)
\item
  Discussion of prolonged persistence. C. difficile sequences detected
  in tissue samples. Association with mucin-degrading bacteria suggested
  by recent papers.
\item
  Discuss why we might not have observed more severe histology in PEG
  mice relative to clindamycin-treated mice

@@

>>>>

\textbf{Five-day post-infection community data can predict mice that
will have prolonged \emph{C. difficile} colonization} After identifying
bacteria associated with the 5-day, 1-day and post-CDI 1-day PEG
treatments, we decided to examine the taxa that were influencing
prolonged \emph{C. difficile} colonization. We trained 3 types of
machine learning models (random forest, logistic regression, and support
vector machine) with input bacterial community data from 5 days
post-infection to predict whether the mice were colonized with \emph{C.
difficile} 10 days post-infection. We chose 5 days post-infection
because that was the earliest timepoint where we would see a treatment
effect in the mice that were given 1-day PEG treatment three days after
\emph{C. difficile} challenge and then administered an FMT or PBS
gavage. The random forest model had the highest performance (AUROC =
0.90, Data Set S1, Sheet X), so we next performed permutation importance
to examine the bacteria that were driving performance. We selected the
top 20 bacteria contributing to our models performance and examined
their relative abundance at 5 days post-infection, the timepoint used to
predict C. difficile colonization status on day 10 (Fig. 6A). Next, we
focused on the 5 genera that were \textgreater{} 1 \% relative abundance
in either the cleared or colonized mice and examined the relative
abundance dynamics of these bacteria over time. We found \emph{, }, and
\_ tended to have a higher relative abundance and \textbf{, } had a
lower relative abundance in the mice with prolonged colonization
compared to the mice that cleared \emph{C. difficile} (Fig. 6B).
Previous work examining the impact of PEG on the murine microbiota found
that PEG treatment resulted in the permanent loss of S24-7, also known
as \emph{Muribaculum intestinale} (16). We decided to check our
\emph{Porphyromonadaceae} OTUs because \emph{Muribaculum intestinale} is
known to be classified as \emph{Porphyromonadaceae} by the Ribosomal
Database Project (RDP) database (22), \emph{Porphyromonadaceae} was a
top feature in the random forest model predicting prolonged \emph{C.
difficile} colonization, and had a high relative abundance in the
communities of mice that cleared \emph{C. difficile} within 10 days. We
identified 4 OTUs that had at least (insert minimum percent identity) to
\emph{Muribaculum intestinale} and examined their abundance in mice that
either cleared or were still colonized with \emph{C. difficile} at 10
days post-infection (Fig. S3).

@@

\begin{itemize}
\item
  Summary of major findings (Fig. 7A)
\item
  Discussion of prolonged persistence. \emph{C. difficile} sequences
  detected in tissue samples. Association with mucin-degrading bacteria
  suggested by recent papers.
\item
  Discuss why we might not have observed more severe histology in PEG
  mice relative to clindamycin-treated mice

@@

\leavevmode\hypertarget{ref-Vornhagen2020}{}%
22. \textbf{Vornhagen J}, \textbf{Bassis CM}, \textbf{Ramakrishnan S},
\textbf{Hein R}, \textbf{Mason S}, \textbf{Bergman Y}, \textbf{Sunshine
N}, \textbf{Fan Y}, \textbf{Timp W}, \textbf{Schatz MC}, \textbf{Young
VB}, \textbf{Simner PJ}, \textbf{Bachman MA}. 2020. A plasmid locus
associated with klebsiella clinical infections encodes a
microbiome-dependent gut fitness factor.
doi:\href{https://doi.org/10.1101/2020.02.26.963322}{10.1101/2020.02.26.963322}.
